Transient Corotating Clumps Around Adolescent Low-Mass Stars From Four Years of TESS

### 2023-09-07 paper/ms.tex
update ms

\subsection{Planets or planetesimal swarms near corotation?}

@@

@@

About one in three young stars with infrared-detected inner dusty
disks show quasiperiodic or stochastic dimming over timescales of
roughly one day \citep[e.g.][]{2010A&A...519A..88A}.  The dimming
amplitudes can reach a few tenths of the stellar brightness, and dips
with identical depths and phases rarely recur.  These ``dipper'' stars
are probably explained by occulting circumstellar dust in the inner
disk
\citep[e.g.][]{2014AJ....147...82C,2016ApJ...816...69A,2021ApJ...908...16R,2022ApJS..263...14C}.
While the phenomenon can persist beyond $\approx$10\,Myr
\citep{2019MNRAS.488.4465G,2022MNRAS.514.1386G}, in all such cases it
seems to be associated with the presence of infrared excesses.
Phenomenologically, dippers are different from CPVs in that their dips
are usually deeper, less periodic, and more variable in depth over
timescales of only one or a few cycles.  Dipper stars also tend to be
younger, since they tend to be classical T Tauri stars with infrared
excesses.

@@

Assuming for the moment that the obscuring material is dust, we can
estimate the mass of a transiting clump. First, we convert the transit
depth into an effective cloud radius, $R_{\rm cloud}$.  For most CPVs
in Figure~\ref{fig:cqvs}, this yields $\approx$2-20\,$R_\oplus$.  A
minimum constraint on the number density of dust particles is obtained
by requiring the cloud to be optically thick.  For cases like LP
12-502, this is reasonable because the transit duration of the
shortest dips implies $R_{\rm cloud}\ll R_\star$.  Carrying out the
relevant calculation assuming the dust grains are 1\,$\mu$m in size,
\citet{2023MNRAS.518.4734S} reported minimum cloud masses of order
$10^{12}$\,kg (their Eq.~23), which scale linearly with both the
optical depth and dust grain radius.  This is comparable to a small
asteroid; the asteroid belt itself has a mass of order
$\approx$10$^{21}$\,kg \citep{2019Icar..319..812P}.  A similar
calculation that assumed occulting clupms of hydrogen, rather than
dust, derived gas prominence masses of at least $10^{14}$\,kg
\citep{1990MNRAS.247..415C}, about 100$\times$ larger than the lower
limit on the dust mass.

>>>>

\subsection{Planets or planetesimals near corotation?}

@@

Assuming for the moment that the obscuring material is dust, we can
estimate the mass of a transiting clump. First, we convert the transit
depth into an effective cloud radius, $R_{\rm cloud}$.  For most CPVs
in Figure~\ref{fig:cqvs}, this yields $\approx$2-20\,$R_\oplus$.  A
minimum constraint on the number density of dust particles is obtained
by requiring the cloud to be optically thick.  For cases like LP
12-502, this is reasonable because the transit duration of the
shortest dips implies $R_{\rm cloud}\ll R_\star$.  Carrying out the
relevant calculation assuming the dust grains are 1\,$\mu$m in size,
\citet{2023MNRAS.518.4734S} reported minimum cloud masses of order
$10^{12}$\,kg (their Eq.~23), which scale linearly with both the
optical depth and dust grain radius.  This is comparable to a small
asteroid; the asteroid belt itself has a mass of order
$\approx$10$^{21}$\,kg \citep{2019Icar..319..812P}.  A similar
calculation that assumed occulting clumps of hydrogen, rather than
dust, derived gas prominence masses of at least $10^{14}$\,kg
\citep{1990MNRAS.247..415C}, about 100$\times$ larger than the lower
limit on the dust mass.

@@

About one in three young stars with infrared-detected inner dusty
disks show quasiperiodic or stochastic dimming over timescales of
roughly one day \citep[e.g.][]{2010A&A...519A..88A,2010ApJS..191..389C}.  The dimming
amplitudes can reach a few tenths of the stellar brightness, and dips
with identical depths and phases rarely recur.  These ``dipper'' stars
are probably explained by occulting circumstellar dust in the inner
disk
\citep[e.g.][]{2014AJ....147...82C,2016ApJ...816...69A,2021ApJ...908...16R,2022ApJS..263...14C}.
While the phenomenon can persist beyond $\approx$10\,Myr
\citep{2019MNRAS.488.4465G,2022MNRAS.514.1386G}, in all such cases it
seems to be associated with the presence of infrared excesses.
Phenomenologically, dippers are different from CPVs in that their dips
are usually deeper, less periodic, and more variable in depth over
timescales of only one or a few cycles.  Dipper stars also tend to be
younger, since they tend to be classical T Tauri stars with infrared
excesses.

@@
